\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}


\begin{document}

%\section{}
%\subsection{}
\begin{center}
{\bf December 7 Meeting}
\end{center}

\begin{itemize}

\item Revise Solutions Manual per Gus' suggestions -- Kipp


\item  Come up with a way to "stamp" the stages of our Schema.
experiment, proposal, recommendation, finalization, v1, v1.1, v1.2, v2, etc. 

\item The (future) supportfor other computational paradigms. I am thinking about robust optimization inparticular, but also about fuzzy optimization and interval arithmetic.

\item Support for Modeling Languages

\begin{itemize}
\item  add support to FlopCpp so that it can generate OSiL (even for nonlinear models)

\item help GAMS people so that GAMS can output OSiL -- ideally work with Stefan so that this can be done through COIN-OR GAMSlinks. -- Wayne


\item convert the LINGO postfix instruction list into a postfix vector of OSnLNodes and then create the OSExpressionTree -- this will allow the LINGO modeling language to directly output OSiL -- Kipp

\item can we do anything with Glpk Mathprog?

\item Other open source modeling languages
\end{itemize}



\item Modify existing OS build system

\begin{itemize}

\item modify the current build so that we can create a lightweight libOS that does not require anything in the Externals except perhaps CppAD -- this will encourage use from developers who don't want to incorporate OS into their software without including a ton of other CoinOR software -- for example, what would be minimal system that GAMS/LINGO developers might want  -- DONE

\item treating examples properly -- create  a Makefile.in for osTestCode so the user can edit a simple Makefile for his or her system.  -- DONE

\end{itemize}
\item  add support for stochastic programming --  modify existing parser to parse a subset of the stochastic nodes  -- Gus and Jun

\item we currently throw an exception if there is an AMPL definitional variable -- treat the AMPL definitional variable correctly  -- Bob

\item use David Gay's idea so that there is no loss of accuracy when writing/reading an OSiL  file  -- Bob


\item talk to Joe Young about Cone programming support -- Joe



\item Project Management and User Support

\begin{itemize}

\item respond to bug reports

\item respond to posts to newsgroups such as Coin-discuss

\item manage release of binaries, stable, and release versions of OS

\item when we need to test who builds what on which platforms?

\end{itemize}




\item Suggestions from Seattle

\begin{itemize}
\item put in the Xpath support so that data is updated in real time

\item support for definitional variables

\item support for a simulation node
\end{itemize}








\end{itemize}


\end{document}  