\division{The CoinEasy Project and CoinAll}

(Much of the material in this section is taken from the CoinEasy web page at\\
{\tt\UrlCoinEasy}.)


As the name implies, the CoinEasy project is designed to make it easy to use COIN-OR projects. It addresses the needs of two groups of users:

\begin{enumerate}

\item Users who want to access COIN-OR solvers to solve optimization problems without having to compile code.

\item Users who want to write applications that use pre-built COIN-OR libraries. 
\end{enumerate}

Users who want to build COIN-OR projects from source code are directed to the home pages of the respective projects for more information.


\ifruncode
This document is aimed specifically at the first group of users. An enlarged document that also includes material for group 2 users can be found at\\
{\tt\UrlCoinAllLibsPdf}.
\else
This document addresses the need of both groups of users. A slimmed-down version that does not include material on the CoinAll libraries can be found at\\
{\tt\UrlCoinAllBinaryPdf}.
\fi

In the information below, we mention the CoinAll and ​CoinBinary  ​ projects.  
CoinAll is  a meta project that consists of most of the solver and utility projects in COIN-OR. 
As such it currently does not have its own web page. 
CoinBinary is a project that provides compiled executable programs and libraries for the projects in CoinAll as well as some other COIN-OR projects. 
Its web page is at {\tt\UrlCoinBinary}. The binary distribution of the CoinAll executables and libraries can be found at {\tt\UrlCoinAllDownload}.

Like other COIN-OR projects, CoinAll has a versioning system that ensures end users some degree of stability 
and a stable upgrade path as project development continues. The current  version of the CoinAll binaries is 1.6.2.

The CoinAll binary distribution includes the following projects

\begin{itemize}
\item{\tt clp - } an open-source linear programming solver \ifuselibs written in C++\fi.

\item{\tt cbc - } an open-source mixed integer programming solver \ifuselibs written in C++\fi.

\item{\tt symphony - } an open-source solver for mixed-integer linear programs (MILPs) \ifuselibs written in C\fi. It supports parallel computations.

\item{\tt ipopt - } a software package designed to find (local) solutions for large-scale ​nonlinear optimization problems.

\item{\tt bonmin - } an experimental open-source C++ code for solving general MINLP (Mixed Integer NonLinear Programming) problems.

\item{\tt couenne - } an exact solver for nonconvex MINLPs.

\item{\tt blis - } a high-performance parallel search implementation for mixed integer linear programs.

\item{\tt OSSolverService - } an integrative framework that allows the other solvers to be called both locally and remotely.

\end{itemize}

OSSolverService is a harness around the other programs, calling any one of them as directed by the user, or as determined by characteristics of the problem.
This document is written mostly from the point of view of explaining and supporting OSSolverService; for information on using the individual solvers in stand-alone form, consult their respective wiki pages.


