\division{File Upload:  Using a File Upload Package}\label{section:fileupload}

\index{OSFileUpload@{\tt OSFileUpload}|(}
When the {\tt OSAgent}\index{OSAgent@{\tt OSAgent}}  class methods {\tt solve}\index{solve@{\tt solve}} and
{\tt send}\index{send@{\tt send}} are used, the problem instance in OSiL\index{OSiL} format is packaged into
a SOAP\index{SOAP protocol} envelope and communication with the server is done using Web Services (for example Tomcat
Axis)\index{Apache Tomcat}. However, packing an XML file into a SOAP envelope may add considerably to the
size of the file (e.g., each {\tt $<$} is replaced with {\tt \&lt;}  and each {\tt $>$} is replaced with {\tt \&gt;}).
Also, communicating with a Web Services servlet can further slow down the communication process.
This could be a problem for large instances. An alternative approach is to use the {\tt OSFileUpload}
executable on the client end and the Java servlet {\tt OSFileUpload} on the server end.  The {\tt OSFileUpload}
client executable is contained in the {\tt fileUpload}  directory inside the {\tt applications} directory.

This servlet is based upon the Apache Commons FileUpload. See 

\medskip
\noindent{\tt\UrlApacheFileupload}
\medskip

The {\tt OSFileUpload} Java class, {\tt OSFileUpload.class} is in the directory
\begin{verbatim}
webapps\os\WEB-INF\classes\org\optimizationservices\oscommon\util
\end{verbatim}
relative to the Web server root.  The source code {\tt OSFileUpload.class} is in the directory
\begin{verbatim}
COIN-OS/OS/applications/fileUpload
\end{verbatim}

Before you can use {\tt OSFileUpload}, you must give a valid URL for the location of the server.
This information must be provided in line 82 of the source code {\tt OSFileUpload.cpp} before 
issuing the {\tt make}\index{make@{\tt make}} command (in a unix environment) or the build (under MS VisualStudio)\index{Microsoft Visual Studio}.

The {\tt OSFileUpload} client executable (see {\tt OS/applications/fileUpload}) takes one argument on the command line,
which is the location of the file on the local directory to upload to the server. For example,
\begin{verbatim}
OSFileUpload ../../data/osilFiles/parincQuadratic.osil
\end{verbatim}
The {\tt OSFileUpload} executable first creates an {\tt OSAgent} object.
\begin{verbatim}
OSSolverAgent* osagent = NULL;
osagent = new OSSolverAgent("http://kipp.chicagobooth.edu/fileupload/servlet/OSFileUpload");
\end{verbatim}
The {\tt OSAgent}  has a method {\tt OSFileUpload} with the signature
\begin{verbatim}
std::string OSFileUpload(std::string osilFileName, std::string osil);
\end{verbatim}
where {\tt osilFileName} is  the name of the OSiL problem instance to be written on the server and {\tt osil}
is the string with the actual instance. Then
\begin{verbatim}
osagent->OSFileUpload(osilFileName, osil);
\end{verbatim}
will place a call to the server, upload the problem instance in the {\tt osil} string, and cause the server
to write on its hard drive a file named {\tt osilFileName}. In our implementation, the uploaded file
({\tt parincQuadratic.osil}) is saved to the {\tt/home/kmartin/temp/parincQuadratic.osil} on the server hard drive.
This location is used in the {\tt osol} file as shown below.

Once the file is on the server, invoke the local {\tt OSSolverService} by
\begin{verbatim}
./OSSolverService config ../data/configFiles/testremote.config
\end{verbatim}
where the {\tt config} file is as follows. Notice there is no {\tt osil}  option as the OSiL file has already
been uploaded and its instance location (``local'' to the server) is specified in the {\tt osol} file.
\begin{verbatim}
osol ../data/osolFiles/remoteSolve2.osol
serviceLocation http://74.94.100.129:8080/OSServer/services/OSSolverService
serviceMethod solve
\end{verbatim}
and the {\tt osol} file is
\begin{verbatim}
<osol xmlns="os.optimizationservices.org"
      xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance"
      xsi:schemaLocation="os.optimizationservices.org
      http://www.optimizationservices.org/schemas/2.0/OSiL.xsd">
    <general>
         <instanceLocation locationType="local">
             /home/kmartin/temp/parincQuadratic.osil
         </instanceLocation>
        <solverToInvoke>ipopt</solverToInvoke>      
    </general>
</osol>
\end{verbatim}

\iffalse   %this needs more work...
As an alternative to using the command line executable {\tt OSFileUpload}, there is also an html form
{\tt fileupload.html} that can be used to upload files. For example, the URL
%\begin{verbatim}
%http://gsbkip.chicagogsb.edu/os/fileupload.html
%\end{verbatim}

\medskip
\noindent{\tt\UrlKippFileupload}
\medskip

\noindent will bring up the necessary form that allows the user to browse a directory and select the file to upload.
This URL is based on the assumption that the {\tt OSJava} classes were deployed as described in
Section~\ref{section:tomcat}. The file {\tt fileupload.html} is in the directory {\tt WebApps/os}.
In our html form implementation, after you upload the OSiL\index{OSiL} file, it shows you the path of the
uploaded file that is saved on the server, so that you can put it in the corresponding {\tt osol} file.
\fi
\index{OSFileUpload@{\tt OSFileUpload}|)}


%\iffalse %------------------------------------------------------
