\division{The  OSInstance API}\label{section:osinstanceAPI}

The OSInstance API can be used to:

\begin{itemize}

\item  get information about model parameters, or convert the {\tt OSExpressionTree} into a prefix or postfix
representation through a collection  of {\tt get()} methods,

\item modify, or even create an instance from scratch, using a number of {\tt set()} methods,

\item provide information to solvers that require function evaluations, Jacobian and Hessian sparsity patters,  
function gradient evaluations, and Hessian evaluations.

\end{itemize}



\subdivision{Get Methods}

The {\tt get()} methods are used by other classes to access data in an existing {\tt OSInstance} object or get 
an expression tree representation of an instance in postfix or prefix format.   Assume {\tt osinstance} is an 
object in the {\tt OSInstance} class created as illustrated in Figure~\ref{figure:creatingosinstanceobject}. 
Then, for example,
\begin{verbatim}
osinstance->getVariableNumber();
\end{verbatim}
will return an integer which is the number of variables in the problem,
\begin{verbatim}
osinstance->getVariableTypes();
\end{verbatim}
will return a {\tt char} pointer to the variable types ({\tt C} for continuous, {\tt B} for binary, 
and {\tt I} for general integer),
\begin{verbatim}
getVariableLowerBounds();
\end{verbatim}
will  return a {\tt double} pointer to the lower bound on each variable. There are similar {\tt get()} methods 
for the constraints. There are numerous {\tt get()} methods for the data in the {\tt <linearConstraintCoefficients>} 
 element, the {\tt <quadraticCoefficients>} element, and the {\tt <nonlinearExpressions>} element.

When an {\tt osinstance} object is created, it is stored as an expression tree in an {\tt OSExpressionTree} object. 
However, some solver APIs (e.g., LINDO) may take the data in a different format such as postfix and prefix. 
There are methods to return the data in either postfix or prefix format.

First define a {\tt vector} of pointers to {\tt OSnLNode} objects.
\begin{verbatim}
std::vector<OSnLNode*> postfixVec;
\end{verbatim}
then get the expression tree for the objective function (index = -1) as a postfix vector of nodes.
\begin{verbatim}
postfixVec = osinstance->getNonlinearExpressionTreeInPostfix( -1);
\end{verbatim}
If, for example, the {\tt osinstance} object was the in-memory representation of   the instance illustrated 
in  Section~\ref{section:rosenbrockXML} and Figure~\ref{figure:expressiontree} then the code
\begin{verbatim}
for (i = 0 ; i < n; i++){
    cout << postfixVec[i]->snodeName << endl;
}
\end{verbatim}
will produce
\begin{verbatim}
number
variable
minus
number
power
number
variable
variable
number
power
minus
number
power
times
plus
\end{verbatim}

This postfix traversal of the expression tree in Figure~\ref{figure:expressiontree} lists all the nodes
by recursively processing all subtrees, followed by the root node.
The method {\tt processNonlinearExpressions()} in the {\tt LindoSolver} class in the {\tt OSSolverInterfaces} 
library component illustrates the use of a postfix vector of {\tt OSnLNode} objects to build a Lindo model instance.


\subdivision{Set Methods}

The {\tt set()} methods can be used to build an in-memory {\tt OSInstance}
 object. A code example of how to do this is in Section~\ref{section:exampleOSInstanceGeneration}.

\subdivision{Calculate Methods}

The {\tt calculate()} methods are described in Section~\ref{section:ad}.


\subdivision{Modifying an   {\tt OSInstance} Object}\label{section:osinstanceMod}

The OSInstance API is designed to be used to either build an in-memory {\tt OSInstance} object 
or provide information about the in-memory object (e.g., the number of variables).   
This interface is not designed for problem modification.  We plan on later providing an {\tt OSModification} 
object for this task. However, by directly accessing an {\tt OSInstance} object it is possible 
to modify parameters in the following classes:

\begin{itemize}
\item {\tt Variables}

\item {\tt Objectives}

\item {\tt Constraints}

\item {\tt LinearConstraintCoefficients}
\end{itemize}

For example, to modify the first nonzero objective function coefficient of the first objective  function to 10.7 the user would write,

\begin{verbatim}
osinstance->instanceData->objectives->obj[0]->coef[0]->value = 10.7;
\end{verbatim}
If the user wanted to modify the actual number of nonzero coefficients as declared by 
\begin{verbatim}
osinstance->instanceData->objectives->obj[0]->numberOfObjCoef;
\end{verbatim}
then the only safe course of action would be to delete the current {\tt OSInstance} object 
and build a new one  with the modified coefficients. It is strongly recommend that no changes 
are made involving allocated memory -- i.e., any kind of {\tt numberOf***}.  
Modifying an objective function coefficient is illustrated in the OSModDemo example. 
See Section \ref{section:exampleOSModDemo}.

After modifying an {\tt OSInstance} object, it is necessary to set certain boolean variables 
to true in order for these changes to get reflected in the OS solver interfaces.

\begin{itemize}
\item {\tt Variables} -- if any changes are made to a parameter in this class set

\begin{verbatim}
osinstance->bVariablesModified = true;
\end{verbatim}

\item {\tt Objectives} -- if any changes are made to a parameter in this class set

\begin{verbatim}
osinstance->bObjectivesModified = true;
\end{verbatim}

\item {\tt Constraints} -- if any changes are made to a parameter in this class set

\begin{verbatim}
osinstance->bConstraintsModified = true;
\end{verbatim}

\item {\tt LinearConstraintCoefficients} -- if any changes are made to a parameter in this class set

\begin{verbatim}
osinstance->bAMatrixModified = true;
\end{verbatim}
\end{itemize}

At this point, if the user desires to modify an {\tt OSInstance} object that contains nonlinear terms, 
the only safe strategy is to delete the object and build a new object that contains the modifications. 



\subdivision{Printing a Model for Debugging}\label{section:printModel}

The OSiL representation for the test problem {\tt rosenbrockmod.osil} is given in 
Appendix~\ref{section:rosenbrockXML}.  Many users will not find the OSiL representation 
useful for model debugging purposes.  For users who wish to see a model in a standard infix 
representation we provide a method {\tt printModel()}.  Assume that we have an {\tt osinstance} 
object in the {\tt OSInstance} class that represents the model of interest.  The call
\begin{verbatim}
osinstance->printModel( -1)
\end{verbatim}
will result in printing the (first) objective function indexed by -1.  In order to print 
constraint~$k$ use
\begin{verbatim}
osinstance->printModel( k)
\end{verbatim}
In order to print the entire model use
\begin{verbatim}
osinstance->printModel( )
\end{verbatim}

 
Below we give the result of {\tt osintance->printModel( )} for the problem {\tt rosenbrockmod.osil}.
\begin{verbatim}
Objectives:
min 9*x_1 + (((1 - x_0) ^ 2) + (100*((x_1 - (x_0 ^ 2)) ^ 2)))

Constraints:
(((((10.5*x_0)*x_0) + ((11.7*x_1)*x_1)) + ((3*x_0)*x_1)) + x_0) <= 25  
10 <= ((ln( (x_0*x_1)) + (7.5*x_0)) + (5.25*x_1))

Variables:
x_0 Type = C  Lower Bound =  0  Upper Bound =  1.7976931348623157e308
x_1 Type = C  Lower Bound =  0  Upper Bound =  1.7976931348623157e308
\end{verbatim}
 
